---
title: "Coweeta Watershed Soil Data Analysis"
author: "D.E. Beaudette"
output: html_notebook
---
# Note: code reformatted by Jonathan Maynard

```{r setup, include=FALSE}
# Load required libraries
library(terra)
library(soilDB)
library(rasterVis)
library(viridisLite)
library(sf)
library(aqp)
library(here)
```

## Input Data:
- Coweeta watershed outlines (UTM z17)
- FY23 RSS grid (NC, EPSG:5070)
- FY23 SSURGO polygons (SDA, WGS84)

## Output Data (UTM z17):
- RSS 10m grid
- SSURGO 10m grid
- SSURGO original polygons
- RSS + SSURGO tabular data as SPC

## Notes:
- "Coweeta_Final_raster.tif" is a non-standard packaging of the RSS
- Grid cells are non-standard keys -> RAT -> [musym] -> mapunit -> [mukey]
    - Use FY23 RSS instead: 10m grid via WCS
    - Tabular RSS data via local download 
- DEM and derivatives will work on a larger area for complete basin characterization

### Coweeta Laboratory Watersheds (outline)
```{r}
# Load Coweeta watershed outline (UTM z17 NAD83)
o <- vect(here('data/vect/Coweeta_Hydrologic_Laboratory.shp'))

# Extend by 200m to ensure a tight crop
b <- buffer(o, width = 200)
```

### Get Latest RSS via WCS
```{r}
# Fetch RSS via WCS, map unit keys as cell values
rss <- mukey.wcs(b, db = 'rss', res = 10)

# Project to UTM z17
rss <- terra::project(rss, crs(o), method = 'near')

# Crop to watershed boundaries + 200m
rss <- crop(rss, b)

# Plot to check
plot(rss)
lines(o)

# Save for later, includes .xml file with RAT
writeRaster(rss, filename = here('data/grids/rss_utm.tif'), overwrite = TRUE)
```

### Get Latest SSURGO via SDA
```{r}
# Get SSURGO polygons for the extended watershed area
s <- SDA_spatialQuery(b, what = 'mupolygon', geomIntersection = TRUE)

# Handle SDA errors
if(inherits(s, 'try-error')) {
  stop('SDA returned an error', call. = FALSE)
}

# Transform to UTM z17
s <- terra::project(s, crs(o))

# Save polygons
writeVector(s, filename = here('data/vect/SSURGO-MU.shp'), overwrite = TRUE)

# Check polygons with RSS
plot(rss)
lines(s)
```

### Rasterize SSURGO Using RSS Grid
```{r}
# Rasterize SSURGO data using RSS grid system
s.rast <- raster::rasterize(s, rss, field = 'mukey')

# Crop and mask to RSS data
s.rast <- mask(crop(s.rast, rss), rss)

# Initialize RAT
s.rast <- as.factor(s.rast)

# Fix names in RAT
set.names(s.rast, 'mukey')
rat <- cats(s.rast)[[1]]
levels(s.rast) <- rat

# Plot to check
plot(s.rast, axes = FALSE)
lines(s)

# Save raster with RAT
writeRaster(s.rast, filename = here('data/grids/ssurgo_utm.tif'), overwrite = TRUE)

aoi_polygon <- st_as_sfc(st_bbox(c(xmin = -83.45, xmax = -83.35, ymin = 35.00, ymax = 35.10), crs = st_crs(s)))
aoi_bbox <- ext(-83.45, -83.35, 35.00, 35.10)
# Define your bounding box coordinates in WGS84 (lat/lon) as a vector
latlon_bbox <- st_bbox(c(xmin = -83.46, xmax = -83.43, ymin = 35.05, ymax = 35.10), crs = st_crs(4326))

# Convert the bbox to an sf polygon for transformation
latlon_polygon <- st_as_sfc(latlon_bbox)

# Transform the bounding box to NAD83 / UTM zone 17N (EPSG:26917)
utm_bbox_polygon <- st_transform(latlon_polygon, crs = 26917)

# Extract the transformed bbox in UTM
utm_bbox <- st_bbox(utm_bbox_polygon)
cropped_line_vector <- crop(s, utm_bbox)
cropped_rast <- crop(s.rast, utm_bbox)
plot(cropped_rast , axes = FALSE)
lines(cropped_line_vector)

```

### Load FY23 RSS Tabular Data
```{r}
# Load local RSS data
.rss_path <- here('data/RSS/RSS_NC.gdb')
st_layers(.rss_path)

# Load relevant tables
rss.mu <- st_read(.rss_path, layer = 'mapunit')
rss.co <- st_read(.rss_path, layer = 'component')
rss.hz <- st_read(.rss_path, layer = 'chorizon')

# Extract RAT from RSS
rat <- cats(rss)[[1]]

# Check for missing mukeys
setdiff(rat$mukey, rss.mu$mukey)
setdiff(rss.mu$mukey, rat$mukey)

# Subset RSS tables
rss.mu <- rss.mu[rss.mu$mukey %in% rat$mukey, ]
rss.co <- rss.co[rss.co$mukey %in% rss.mu$mukey, ]
rss.hz <- rss.hz[rss.hz$cokey %in% rss.co$cokey, ]

# Save raw tabular data
save(rss.mu, rss.co, rss.hz, file = here('data/rss-tab-data-raw.rda'))
```

### Load FY23 SSURGO Tabular Data
```{r}
# Extract mukeys from rasterized SSURGO data
rat <- cats(s.rast)[[1]]
.mukeys <- as.numeric(rat$mukey)

# Query SSURGO tabular data via SDA
ssurgo.mu <- SDA_query(sprintf("SELECT * FROM mapunit WHERE mukey IN %s", format_SQL_in_statement(.mukeys)))
ssurgo.co <- SDA_query(sprintf("SELECT * FROM component WHERE mukey IN %s", format_SQL_in_statement(.mukeys)))

# Get component keys and query horizons
.cokeys <- unique(ssurgo.co$cokey)
ssurgo.hz <- SDA_query(sprintf("SELECT * FROM chorizon WHERE cokey IN %s", format_SQL_in_statement(.cokeys)))

# Save SSURGO tabular data
save(ssurgo.mu, ssurgo.co, ssurgo.hz, file = here('data/ssurgo-tab-data-raw.rda'))
```

### Combine and Analyze RSS and SSURGO Data
```{r}
# Upgrade tabular data to SoilProfileCollection
depths(rss.hz) <- cokey ~ hzdept_r + hzdepb_r
site(rss.hz) <- rss.co
site(rss.hz) <- rss.mu

depths(ssurgo.hz) <- cokey ~ hzdept_r + hzdepb_r
site(ssurgo.hz) <- ssurgo.co
site(ssurgo.hz) <- ssurgo.mu

# Plot RSS and SSURGO soil profile collections
plotSPC(rss.hz[1:10, ])
plotSPC(ssurgo.hz[1:10, ])

# Combine RSS and SSURGO data
site(rss.hz)$source <- 'RSS'
site(ssurgo.hz)$source <- 'SSURGO'
spc <- c(rss.hz, ssurgo.hz)

# Set horizon name
hzdesgnname(spc) <- 'hzname'

# Check the number of profiles
length(spc)

# Compute depth class
sdc <- getSoilDepthClass(spc)
site(spc) <- sdc

# Summarize data
table(spc$depth.class)
table(spc$source, spc$depth.class)

# Save combined tabular data as SPC
saveRDS(spc, file = here('data/combined-tab-data-SPC.rds'))
```

