---
title: "Soil Data Aggregation and Analysis"
author: "Dylan Beaudette"
output: html_notebook
---

# Note: code reformatted by Jonathan Maynard

```{r setup, include=FALSE}
# Load necessary libraries
library(aqp)
library(lattice)
library(tactile)
library(reshape2)
library(here)
# Load local functions
source(here('code/local-functions.R'))
```

## TODO: Think about the implications of aggregation 
- Depth class -> texture class.
- Does it make sense to do these separately?

### Load Data
```{r}
# Load data from '01-prepare-soil-data.R'
# Combined map unit/component data as SPC (SoilProfileCollection)
x <- readRDS(here('data/combined-tab-data-SPC.rds'))

# Investigate mukind by source
xtabs(~ source + mukind, data = unique(site(x)[, c('mukey', 'source', 'mukind')]))
```

### Original Proportions of Depth Class by Source
```{r}
# Analyze depth class proportions by source
xtabs(~ source + depth.class, data = site(x))

# Visualize depth class distribution by source
histogram(~ depth | source, data = site(x), par.settings = tactile.theme())
```

### Simplify Depth Classes
```{r}
# Classify shallow (<50cm) vs everything else
x$depth.class <- factor(x$depth.class, ordered = TRUE)
x$simple.depth.class <- ifelse(x$depth.class <= 'shallow', 'shallow', 'deep')

# Check proportions of simplified depth classes
prop.table(xtabs(~ source + simple.depth.class, data = site(x)), margin = 1)

# Develop dominant condition: simplified soil depth class
s <- site(x)
z <- split(s, s$mukey)
z <- lapply(z, dominantCondition, v = 'simple.depth.class')

# Create LUT (Lookup Table) by mukey/cokey
depth.lut <- do.call('rbind', z)
```

### Analyze Soil Texture Class by Depth Condition
```{r}
# Filter SPC based on cokey in depth LUT
x.sub <- subset(x, cokey %in% unique(depth.lut$cokey))

# Classify soil texture (<2mm)
x.sub$texture <- ssc_to_texcl(sand = x.sub$sandtotal_r, clay = x.sub$claytotal_r)

# Visualize soil profiles with texture classification
par(mar = c(0, 0, 3, 1))
plotSPC(x.sub[1:30, ], color = 'texture')
plotSPC(x.sub, color = 'texture', name = '', print.id = FALSE, width = 0.35)
```

## TODO: Is this a reasonable approach?
- Consider 0-50cm depth.
- Derive depth-weighted mean soil texture (sand and clay content).

### Compute Depth-Weighted Mean Soil Texture
```{r}
# Compute weighted mean sand and clay content to depth (0-200cm)
a <- slab(x.sub, cokey ~ sandtotal_r + claytotal_r, slab.structure = c(0, 200), slab.fun = mean, na.rm = TRUE)

# Reshape and classify soil texture
w <- dcast(a, cokey ~ variable, value.var = 'value')
w$texture <- ssc_to_texcl(sand = w$sandtotal_r, clay = w$claytotal_r)
```

### Check Degree of Aggregation
```{r}
# Merge original and aggregated texture data
g <- merge(horizons(x.sub)[, c('cokey', 'texture')], w[, c('cokey', 'texture')], by = 'cokey', all.x = TRUE, sort = FALSE)
names(g) <- c('cokey', 'original.texture', 'agg.texture')
g$agg.texture <- factor(g$agg.texture, levels = levels(g$original.texture))

# Cross-tabulate original vs aggregated texture
g.tab <- table(original = g$original.texture, aggregated = g$agg.texture)
round(prop.table(g.tab, margin = 1) * 100)
```

## TODO: Report on the Loss of Detail
- Examine the loss of detail, particularly abrupt textural changes.
```{r}
# Visualize original texture and aggregated texture
plotSPC(x.sub[1:5, ], color = 'texture')
w[1:5, ]
```

### Create Final LUT and Define "Soil Type"
```{r}
# Merge depth LUT and texture data
final.lut <- merge(depth.lut, w, by = 'cokey', all.x = TRUE, sort = FALSE)

# Define soil type as interaction between depth class and texture
final.lut$soil.type <- interaction(final.lut$simple.depth.class, final.lut$texture)

# Check final soil type table by source
table(final.lut$source, final.lut$soil.type)

# Save final LUT
saveRDS(final.lut, file = here('data/soil-depth-texture-classe-lut.rds'))
```
