---
title: "Create Soil Maps"
author: "D.E. Beaudette"
output: html_notebook
---

# Note: code reformatted by Jonathan Maynard

```{r setup, include=FALSE}
# Load necessary libraries
library(terra)
library(aqp)
library(viridisLite)
library(here)
```

## Loading 10m Mukey Grids (UTM z17)
```{r}
# Load mukey grids
r <- rast(here('data/grids/rss_utm.tif'))
s <- rast(here('data/grids/ssurgo_utm.tif'))
```

## Load Combined LUT with Depth/Texture Classes
```{r}
# Load LUT with depth/texture classes and soil type class
lut <- readRDS(here('data/soil-depth-texture-classe-lut.rds'))
```

### Check on Codes Used by Source
```{r}
# Check structure of soil type classes
str(lut$soil.type)

# List soil type levels
levels(lut$soil.type)

# Frequency tables by source and soil type
table(source = lut$source, soil.type = lut$soil.type)

# Frequency tables by source and component name
table(source = lut$source, compname = lut$compname)
```

## Ensure Factor Levels and Order are Preserved
```{r}
# Convert soil type to numeric codes
lut$soil.type.numeric <- as.numeric(lut$soil.type)
```

## Extract and Update Raster Attribute Tables (RAT)
```{r}
# Extract RAT from RSS and SSURGO grids
r.rat <- cats(r)[[1]]
s.rat <- cats(s)[[1]]

# Merge with LUT
r.rat <- merge(r.rat, lut, by = 'mukey', all.x = TRUE, sort = FALSE)
s.rat <- merge(s.rat, lut, by = 'mukey', all.x = TRUE, sort = FALSE)

# Convert 'mukey' to numeric
r.rat$mukey <- as.numeric(r.rat$mukey)
s.rat$mukey <- as.numeric(s.rat$mukey)

# Re-pack updated RATs
levels(r) <- r.rat
levels(s) <- s.rat
```

## Convert RAT Columns to Raster Layers
```{r}
# Convert RAT columns to raster layers
rr <- catalyze(r)
ss <- catalyze(s)

# Plot component names
plot(rr['compname'])
plot(ss['compname'])
```

## Combine and Plot Grids
```{r}
# Combine RSS and SSURGO grids based on soil type numeric codes
z <- c(rr['soil.type.numeric'], ss['soil.type.numeric'])

# Plot combined grids with rainbow colors
plot(z, col=rainbow(5))
```

## Save Soil Type Class Rasters
```{r}
# Save RSS and SSURGO soil type class rasters
writeRaster(rr['soil.type.numeric'], filename = here('data/grids/rss-soiltype-class.tif'), overwrite = TRUE)
writeRaster(ss['soil.type.numeric'], filename = here('data/grids/ssurgo-soiltype-class.tif'), overwrite = TRUE)
```

## Plot Saved Grids for Verification
```{r}
col=rainbow(5)
# Load and plot saved soil type rasters
plot(rast(here('data/grids/rss-soiltype-class.tif')), col=col[-4])
plot(rast(here('data/grids/ssurgo-soiltype-class.tif')), col=col)
```

## Optional: Testing Soil Type Integer to Factor Conversion (Not Working)
```{r eval=FALSE}
# Testing: convert soil type integers to factors (doesn't work)
# z <- as.factor(z)
# ll <- cats(z)
# 
# .st <- levels(lut$soil.type)
# ll[[1]]$soiltype <- .st[ll[[1]]$label]
# ll[[2]]$soiltype <- .st[ll[[2]]$label]
# 
# levels(z) <- ll
# 
# activeCat(z[[1]]) <- 'soiltype'
# activeCat(z[[2]]) <- 'soiltype'
# plot(z)
```

